---
output: github_document
---

<!-- README.md is generated from README.Rmd — please edit that file, then `npm run readme:rmd`. -->
<!-- The `pi`, `dogfood`, and `biocli` chunks run real commands. Rendering needs a built `dist/`, Pi, and model credentials. -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(comment = "")

wrap_prompt <- function(p, width = 56) {
  words <- strsplit(p, " ")[[1]]
  lines <- character(0)
  cur <- ""
  for (w in words) {
    if (nzchar(cur) && nchar(cur) + nchar(w) + 1 > width) {
      lines <- c(lines, cur)
      cur <- w
    } else {
      cur <- if (nzchar(cur)) paste(cur, w) else w
    }
  }
  lines <- c(lines, cur)
  paste(vapply(seq_along(lines), function(i) {
    paste0('  "', lines[i], if (i < length(lines)) " " else "", '"', if (i < length(lines)) " \\" else "")
  }, ""), collapse = "\n")
}

sanitize_paths <- function(txt) {
  wd <- normalizePath(getwd(), winslash = "/", mustWork = FALSE)
  txt <- gsub(paste0(wd, "/"), "", txt, fixed = TRUE)
  txt <- gsub(wd, ".", txt, fixed = TRUE)
  txt <- gsub("rare_high_impact\\.report-[0-9]+-[0-9a-f]+", "rare_high_impact.report-<run>", txt, perl = TRUE)
  txt <- gsub("query-[0-9]+-[0-9a-f]+", "query-<run>", txt, perl = TRUE)
  txt <- gsub("reproduce-readme-variant-counts-[0-9]+", "reproduce-readme-variant-counts-<run>", txt, perl = TRUE)
  txt <- gsub("pid[ =][0-9]+", "pid <pid>", txt, perl = TRUE)
  txt <- gsub("/tmp/pi-bio-handoff-cas-[0-9]+", "/tmp/pi-bio-handoff-cas-<run>", txt, perl = TRUE)
  txt <- gsub("[ \t]+(?=\n|$)", "", txt, perl = TRUE)
  txt
}

knitr::knit_engines$set(pi = function(options) {
  prompt <- paste(options$code, collapse = " ")
  model <- if (!is.null(options$model)) options$model else "openai-codex/gpt-5.3-codex"
  thinking <- if (!is.null(options$thinking)) options$thinking else "high"
  tools <- if (!is.null(options$tools)) options$tools else "read,write,bio_describe_model,bio_query,bio_list_duckdb_extensions"
  tmo <- if (!is.null(options$timeout)) options$timeout else 300
  must_contain <- if (!is.null(options$mustContain)) as.character(options$mustContain) else character(0)
  dir.create(".pi/bio-agent", recursive = TRUE, showWarnings = FALSE)
  if (!is.null(options$removeBefore)) unlink(options$removeBefore, force = TRUE)
  args <- c("--model", model, "--thinking", thinking, "--no-extensions", "-e", "extensions/pi-coding-agent/index.ts", "--tools", tools, "-p", "--no-session", prompt)
  out <- suppressWarnings(system2("pi", shQuote(args), stdout = TRUE, stderr = TRUE, timeout = tmo))
  status <- attr(out, "status")
  out_txt <- sanitize_paths(paste(out, collapse = "\n"))
  if (!is.null(status) && status != 0) stop("pi render failed (exit ", status, "):\n", out_txt)
  for (needle in must_contain) {
    if (nzchar(needle) && !grepl(needle, out_txt, fixed = TRUE)) stop("pi render did not produce required output '", needle, "':\n", out_txt)
  }
  knitr::asis_output(paste0(
    "``` sh\npi \\\n",
    "  --model ", model, " \\\n",
    "  --thinking ", thinking, " \\\n",
    "  --no-extensions \\\n",
    "  -e extensions/pi-coding-agent/index.ts \\\n",
    "  --tools ", tools, " \\\n",
    "  -p --no-session \\\n",
    wrap_prompt(prompt), "\n```\n\n",
    out_txt, "\n"
  ))
})

knitr::knit_engines$set(biocli = function(options) {
  args <- paste(options$code, collapse = " ")
  out <- suppressWarnings(system(paste("node dist/cli/bin.js", args, "2>&1"), intern = TRUE))
  status <- attr(out, "status")
  out_txt <- sanitize_paths(paste(out, collapse = "\n"))
  if (!is.null(status) && status != 0) stop("biocli render failed (exit ", status, "): ", args, "\n", out_txt)
  lang <- if (grepl("^\\s*[\\[{]", out_txt)) "json" else ""
  # wrap each flag onto its own continuation line so long commands don't overflow the README
  cmd <- gsub(" --", " \\\\\n  --", paste("pi-bio-agent", args))
  knitr::asis_output(paste0("``` sh\n", cmd, "\n```\n\n", "``` ", lang, "\n", out_txt, "\n```\n"))
})

knitr::knit_engines$set(dogfood = function(options) {
  cmd <- paste(options$code, collapse = " ")
  out <- suppressWarnings(system(paste(cmd, "2>&1"), intern = TRUE))
  status <- attr(out, "status")
  out_txt <- sanitize_paths(paste(out, collapse = "\n"))
  if (!is.null(status) && status != 0) stop("dogfood render failed (exit ", status, "): ", cmd, "\n", out_txt)
  start <- regexpr("\\{\\n  \"dogfood\"", out_txt)
  if (start < 0) stop("dogfood render did not print summary JSON:\n", out_txt)
  summary <- jsonlite::fromJSON(substr(out_txt, start, nchar(out_txt)), simplifyVector = FALSE)
  artifact_roles <- Filter(function(x) !is.null(x$semanticRole), summary$trainingCorpus$artifactRoles)
  artifact_roles <- lapply(artifact_roles, function(x) Filter(Negate(is.null), x))
  proof <- list(
    observationCounts = summary$observationCounts[c("executes", "invoked_by", "tool_call", "run", "host_event", "job_step_checkpoint")],
    ducknngProfile = summary$ducknngProfileReceipt[c("version", "receiptChanged", "subjectRestriction")],
    casMetadata = summary$hostCapabilityRun[c("casMetadataRefs", "artifactCasMetadataRefs")],
    graphProjection = list(
      external = summary$externalProjection[c("edgeCount", "closureCount")],
      internal = list(
        edgeCountAtLeast4 = summary$internalProjection$edgeCount >= 4,
        closureCountAtLeast3 = summary$internalProjection$closureCount >= 3
      )
    ),
    trainingCorpus = c(
      summary$trainingCorpus[c("toolCalls", "runs", "artifacts", "hostEvents", "hostEventLinks", "parquetReadbackRows")],
      list(artifactRoles = artifact_roles)
    ),
    governanceHostEvents = summary$governanceHostEvents[c("count", "linkCount", "kinds", "eventTypes")],
    sdkConsumer = summary$sdkConsumer[c("publicExportsOnly", "runtimeImports", "packageSource")],
    queueCancel = summary$queueCancel[c("staleWriteRejected", "resumedPhase")],
    renvEnvironment = summary$renvEnvironment[c("rVersion", "bioconductor", "envStatus", "artifactRows")]
  )
  knitr::asis_output(paste0("``` sh\n", cmd, "\n```\n\n``` json\n", jsonlite::toJSON(proof, pretty = TRUE, auto_unbox = TRUE), "\n```\n"))
})
```

# pi-bio-agent

[![CI](https://github.com/sounkou-bioinfo/pi-bio-agent/actions/workflows/ci.yml/badge.svg)](https://github.com/sounkou-bioinfo/pi-bio-agent/actions/workflows/ci.yml)
[![License: GPL v2+](https://img.shields.io/badge/License-GPL%20v2%2B-blue.svg)](LICENSE)
[![Node](https://img.shields.io/badge/node-%3E%3D20-brightgreen.svg)](package.json)

`pi-bio-agent` is a provider-agnostic bioinformatics agent substrate and Pi extension
for agent-controlled scientific computation: manifests + SQL over DuckDB are the
program. A manifest declares resources, read-only SQL, optional process compute,
receipts, CAS artifacts, and observations in one DuckDB-backed ledger.

The agent may inspect schemas and write SQL; biomedical facts come from declared
data, deterministic compute, receipts, and recorded approvals.

## Bets

- New biomedical questions should become manifests, SQL, operation specs, or graph
  data, not new per-question tools.
- DuckDB is the substrate: files, formats, HTTP-shaped data, graph tables, and
  reductions should be reached through SQL and extensions whenever possible.
- Knowledge belongs in queryable graph tables. `bio_edges`, `bio_edges_as_of`, and
  `entailed_edge` are the interaction surface, not serialized graph context.
- CAS, resolver receipts, replay specs, and environment evidence are part of the
  result, not optional metadata.
- Compute is a host-injected async port: submit, status, collect, cancel. The core
  records the contract, durable queue status, step checkpoints, replay specs, and
  evidence; hosts decide what processes, credentials, and network access are
  allowed. Resume is checkpoint-based, not a workflow engine.
- The model or human handles routing, schema inspection, SQL composition, and
  typed judgment over an apparatus of manifests, tables, graphs, receipts, and
  gates. It is not the source of biomedical facts; constraints are what make the
  computation auditable.

## See It

The agent starts with the raw ClinVar VCF URL and a TP53 genomic range, writes the
manifest into `.pi/`, runs the query, and returns the SQL plus results. This is a
tabix range read through `duckhts`, so it should complete in seconds rather than
download the whole ClinVar VCF. ClinVar is live data; the rows below are the
result from this README render, not a pinned truth table.

```{pi, timeout=300, removeBefore=".pi/bio-agent/readme-clinvar-tp53.json", mustContain=c("Pathogenic", "GROUP BY", "UNNEST", "|")}
Create .pi/bio-agent/readme-clinvar-tp53.json for the raw ClinVar VCF URL https://ftp.ncbi.nlm.nih.gov/pub/clinvar/vcf_GRCh38/clinvar.vcf.gz, region 17:43044295-43125483, table clinvar, using duckhts.read_bcf. Run bio_query to count clinical-significance buckets in that region. Inspect the schema if needed; avoid double-counting multi-valued significance entries. Return only the manifest path, the SQL, and a Markdown result table.
```

The same query with no model in the loop:

```{biocli}
query .pi/bio-agent/readme-clinvar-tp53.json --db :memory: --init-sql "INSTALL duckhts FROM community; LOAD duckhts;" --sql "WITH distinct_calls AS (SELECT DISTINCT CHROM, POS, REF, ALT, significance FROM clinvar, UNNEST(INFO_CLNSIG) AS u(significance)) SELECT significance, COUNT(*) AS n FROM distinct_calls GROUP BY significance ORDER BY n DESC"
```

## Start

Install the Pi extension:

```sh
pi install git:github.com/sounkou-bioinfo/pi-bio-agent
/reload
```

Use the CLI without Pi:

```{r cli-ledger-reset, include=FALSE}
unlink(c(".pi/bio-agent/readme-store.duckdb", ".pi/bio-agent/readme-store.duckdb.wal"), force = TRUE)
```

```{biocli}
query examples/variant-counts/manifest.json --db :memory: --run-id readme-variant-counts --ledger .pi/bio-agent/readme-store.duckdb --author readme --sql "SELECT consequence, count(*) AS n FROM variants GROUP BY consequence ORDER BY consequence"
```

Verify that run against a fresh database:

```{biocli}
reproduce .pi/bio-agent/runs/readme-variant-counts/replay.json
```

Use it as a library:

```ts
import { runBioQueryFromManifest } from "pi-bio-agent";

const out = await runBioQueryFromManifest({
  cwd: process.cwd(),
  dbPath: ":memory:",
  manifestPath: "examples/variant-counts/manifest.json",
  sql: "SELECT consequence, count(*) AS n FROM variants GROUP BY consequence",
});
```

Run the full gate:

```sh
npm install
npm run check
```

`npm run check` runs typecheck, tests, docs/example drift checks, skill conformance, and README tool-list checks. CI
provisions real R, pinned `nanoarrow`, and DuckHTS before the full suite. A separate owned-ducknng job requires
retry, subject-scoped HTTP profiles, upload, TLS, RPC, and socket behavior.

## How it works

A manifest declares named resources. Resolvers materialize them into DuckDB tables and
stamp receipts: resolver version, parameter digest, and source snapshot. An operation
is one read-only `SELECT`/`WITH` over those tables. Whatever it returns is the result.

| leg | primitive |
|---|---|
| Data | `file_scan` and `duckdb.sql_materialize` over files, object stores, local DBs, and DuckDB community extensions such as `duckhts`, `anndata`, `duckdb_zarr`, and `plinking_duck`. |
| Network | `ducknng_ncurl_table` and ducknng HTTP profiles for SQL-native HTTP/RPC. `http.get` remains a host-injected JS fetch resolver when an application chooses it. |
| Compute | `compute.run`: a child process over Arrow IPC and declared file artifacts. R, Python, `rv`, Nextflow, or an NNG worker pool are argv/runner choices, not new core APIs. |
| Knowledge | one SemanticSQL-shaped graph: `bio_edges`, `bio_edges_as_of`, and `entailed_edge` for ontology terms, memory, observations, sessions, tool calls, artifacts, and run links. |

Host capabilities are explicit. If no `fetch` is injected, `http.get` is unavailable.
If no `ComputeRunner` is injected, `compute.run` is unavailable. The library validates
manifests, SQL shape, receipts, and replay. The host owns filesystem, network,
credentials, and process isolation.

The plain CLI keeps those grants explicit: `--network fetch` binds a capped WHATWG-fetch adapter,
`--compute local` binds the local async runner, `--cas-root` captures outputs, and host-owned config/protected values
come from JSON files rather than manifest data. An embedding host can replace each port with stricter policy.

Manifest inspection does not infer runnability from an operation's transport. It reports declarations separately
from this host's `ready`, `blocked`, or `unknown` admission status. A bound resolver with an unattested extension or
egress requirement remains `unknown` until the host proves it.

For credentialed SQL-native HTTP, the CLI can commission a ducknng HTTP profile on the
same DuckDB connection as the run. The profile file contains non-secret policy plus a
credential source; the token comes from env or stdin, and replay records only the
redacted profile receipt digest.

```json
{
  "profileId": "clinvar-read",
  "scheme": "https",
  "host": "api.example.org",
  "pathPrefix": "/clinvar",
  "method": "GET",
  "tlsRequired": true,
  "authHeaderName": "Authorization",
  "authHeaderValueEnv": "CLINVAR_TOKEN",
  "allowSubjects": ["case:alpha"]
}
```

```sh
pi-bio-agent query credentialed-manifest.json \
  --db :memory: \
  --init-sql "LOAD ducknng" \
  --ducknng-http-profile ./clinvar-profile.json \
  --sql "SELECT * FROM secured_table LIMIT 5"
```

`credentialed-manifest.json` is the host/app manifest whose `ducknng_ncurl_table` call names the profile id
`clinvar-read`; `CLINVAR_TOKEN` is supplied by the host environment or secret manager, and the secret never appears
in the manifest, SQL, or argv.

## Ledger

Runs, facts, memory, job status, session traces, and artifacts are rows in
`bio_observations`, read as of a time. The graph view is derived from edge-like
observations, so these are ordinary SQL questions:

- what is the current fact?
- what did the agent learn?
- what was this job's status at a time?
- which tool call produced this scientific run?
- which manifest, operation, and resources did it use?

When a host supplies the ledger, run facts, manifest-declaration links, and declared artifact projections are
required evidence. A recording failure is returned with the already-persisted run path; it is not silently converted
into success. `pi-bio-agent reproduce` and `bio_reproduce_run` re-execute `replay.json` against a fresh database and
report source, output, or environment drift.

Persisted Pi session JSONL and Codex rollout JSONL are ingested as `session:`, `turn:`, `msg:`,
`toolcall:`, and `cas:` observations through the same ledger contract. Pi's extension syncs its active session;
any host can use the CLI importer:

```sh
pi-bio-agent session import <session.jsonl> --format pi
pi-bio-agent session import <rollout.jsonl> --format codex
```

The original JSONL is streamed into CAS and normalization reads that immutable snapshot, so a live file growing
during import cannot disagree with its digest. Large traces commit bounded, idempotent batches; the terminal
`session` fact is the completion marker and a retry does not duplicate prior statements. Persisted transcripts
recover messages, tool trajectories, compaction, and parentage. Runtime-only delivery/interrupt signals still need
a host hook. Bio tools record controlled run links at execution time; the ingester never scans transcript text for
run-looking strings.

Window an edge-shaped graph table without loading the whole neighborhood:

```{r graph-window-dogfood-reset, include=FALSE}
dir.create(".pi/bio-agent", recursive = TRUE, showWarnings = FALSE)
unlink(c(".pi/bio-agent/readme-graph-window.duckdb", ".pi/bio-agent/readme-graph-window.duckdb.wal"), force = TRUE)
```

```{biocli}
query examples/graph-window/manifest.json --db .pi/bio-agent/readme-graph-window.duckdb --sql "SELECT count(*) AS n FROM bio_edges"
```

```{biocli}
graph-window --db .pi/bio-agent/readme-graph-window.duckdb --table bio_edges --start run:readme --direction both --limit 10
```

## Dogfood

Run the compact all-primitives dogfood:

```{dogfood}
npm run dogfood:bring-it-home
```

It exercises host-event receipts, step checkpoints/resume, SemanticSQL projection,
ducknng profile receipts, SDK exports, figure/report CAS artifacts, and real process
compute when R is available. SemanticSQL compatibility is pinned to the concrete generated-view names at upstream
commit `83503077e867419c18a211d97105d8ead554e947`; it is not a claim of equivalence with every future generator.

Run the Pi session trace dogfood after installing the extension and configuring an
image-capable model:

```sh
npm run dogfood:pi-session-trace
```

It drives a real Pi session through image read, successful shell call, intentional
shell error, manifest discovery, SQL validation, `bio_query`, and then queries the
project ledger for the session summary, tool trajectory, CAS roots, and the recorded
tool-call/run linkage.

## Examples

Every [example](examples/) has a recorded run; `npm run check` fails if generated
example blocks drift.

- [`examples/variant-counts`](examples/variant-counts/) - ad-hoc SQL over a CSV resource.
- [`examples/rare-high-impact`](examples/rare-high-impact/) - a declared operation with explicit abstention.
- [`examples/compute-run`](examples/compute-run/) - DuckDB table -> Arrow IPC -> R `lm()` -> DuckDB table.
- [`examples/compute-artifacts`](examples/compute-artifacts/) - compute value plus declared files captured into CAS.
- [`examples/compute-files-only`](examples/compute-files-only/) - a file-only tool exposed as an artifacts table.
- [`examples/coloc`](examples/coloc/) - post-GWAS colocalization: SQL harmonization plus out-of-process R.
- [`examples/connectors`](examples/connectors/) - APIs as manifests, including UniProt, RCSB PDB, MyGene, Reactome, Monarch, and OpenTargets GraphQL.

Example connector, no TypeScript client:

```{biocli, cache=TRUE}
query examples/connectors/uniprot.json --db :memory: --init-sql "LOAD ducknng; SET VARIABLE tls = ducknng_tls_config_from_pem(NULL, NULL, NULL, '', 1)" --bindings '{"uniprot_acc":"P04637"}' --sql "SELECT primaryAccession, uniProtkbId, sequence.length aa FROM uniprot_entry"
```

## Pi tools

The `pi-coding-agent` extension registers these tools over the substrate. This list is generated from the
extension's `registerTool()` calls (`npm run readme:tools`); `npm run check` fails if it drifts.

<!-- BEGIN GENERATED:tools (scripts/generate-readme-tools.mjs — do not edit by hand) -->

- `bio_list_sources` — List manifest-backed sources
- `bio_describe_model` — Describe Pi Bio model
- `bio_run_operation` — Run a bio operation
- `bio_query` — Run an ad-hoc bio query
- `bio_reproduce_run` — Reproduce a bio run
- `bio_list_duckdb_extensions` — List bio DuckDB extensions
- `bio_validate_select` — Validate bio SQL SELECT
- `bio_validate_graph_projection` — Validate graph projection
- `bio_graph_window` — Window graph context
- `bio_create_skill` — Create bio skill
- `bio_study_plan` — Plan bio study
- `bio_remember` — Remember (memory note)
- `bio_list_memory` — List memory
- `bio_walk_memory` — Walk bio memory graph
- `bio_recall` — Recall memory note
- `bio_forget` — Forget memory note

<!-- END GENERATED:tools -->

Project-local skills and the memory store live under `.pi/bio-agent/` in the current project.

## Host-neutral agent skill

The package also ships `skills/pi-bio-agent/`: a procedural guide for hosts that do not run the Pi extension. It is
not a biomedical computation pack. It tells an agent to discover and describe manifests, inspect tables with
`DESCRIBE` / `SUMMARIZE`, run read-only SQL through `pi-bio-agent query`, and use `graph-window` for bounded ledger or
KG walks.

Pi users should normally install the full package. That gives Pi the `bio_*` tools, the plain skill, and session trace
integration that links tool calls to the recorded `run:<id>` they produced:

```sh
pi install git:github.com/sounkou-bioinfo/pi-bio-agent
/reload
```

Install only the skill, directly from GitHub and without cloning, when another host should call the CLI:

```sh
npx --yes github:sounkou-bioinfo/pi-bio-agent install-skill --host codex
npx --yes github:sounkou-bioinfo/pi-bio-agent install-skill --host pi-project
npx --yes github:sounkou-bioinfo/pi-bio-agent install-skill --host claude-project
npx --yes github:sounkou-bioinfo/pi-bio-agent install-skill --host opencode-project
npx --yes github:sounkou-bioinfo/pi-bio-agent install-skill --host copilot-project
npx --yes github:sounkou-bioinfo/pi-bio-agent install-skill --dest /path/to/host/skills
```

Supported `--host` presets include `codex`, `pi`, `pi-project`, `claude`, `claude-project`, `opencode`,
`opencode-project`, `copilot`, and `copilot-project`. Use `--dest` for any host-specific skill directory not covered
by a preset.

The skill install does not guarantee the CLI is on `PATH`. Install the CLI persistently when the host should run
`pi-bio-agent query` / `pi-bio-agent run` later:

```sh
npm install -g github:sounkou-bioinfo/pi-bio-agent
pi-bio-agent install-skill --dest /path/to/host/skills
```

From a checkout, install the skill for local development and optionally link the CLI:

```sh
npm install
npm run build
npm run install:skill -- --dest /path/to/host/skills
npm run install:skill -- --dest /path/to/host/skills --link-cli
```

Run `npx --yes github:sounkou-bioinfo/pi-bio-agent install-skill --help` for all presets and flags. The skill is
guidance; manifests, SQL, receipts, CAS, and observations remain the computation.

## Docs

Start with the [user guide](docs/guide.md): write a manifest, run an operation. Then
read [design notes](docs/design.md) for the core contracts and [roadmap](docs/roadmap.md)
for current work. The full [docs index](docs/INDEX.md) is generated from frontmatter.

## References & lineage

Prior art and lineage:

- **ClawBio**, the origin corpus this factors into manifests, resolvers, and operations:
  <https://github.com/ClawBio/ClawBio>
- **metacurator**, deterministic curation stages plus typed model judgments:
  <https://github.com/seandavi/metacurator>
- **op2workshop / ontoProc2**, the workshop lineage that led us back to the Semantic SQL source spec:
  <https://github.com/vjcitn/op2workshop>
- **Machine studying** (Li, Battle, Khattab, 2026): <https://jacobxli.com/blog/2026/machine-studying/>
- **Sakana Fugu** (learned orchestration over shared memory and access lists): <https://sakana.ai/fugu/>
- **Recursive Language Models / RLM** (symbolic code and recursive model calls over external context; `bio_query`
  supplies the relational data-plane part, not the recursive model loop): <https://arxiv.org/abs/2512.24601>
- **ducknng**, an Arrow-native DuckDB extension for NNG/HTTP/RPC transport, in the lineage of R's
  `nanonext` + `mirai`: <https://github.com/sounkou-bioinfo/ducknng> · [NNG](https://nng.nanomsg.org/) ·
  [`nanonext`](https://github.com/r-lib/nanonext) · [`mirai`](https://mirai.r-lib.org/)
- **SemanticSQL** (canonical LinkML source schema for `statements`, `prefix`, generated views, and
  `entailed_edge`): <https://github.com/INCATools/semantic-sql>
- Design thread (sounkou-bioinfo × Manuel): [LinkedIn](https://www.linkedin.com/feed/update/urn:li:activity:7473824764575436800)

## Development

```sh
npm install
npm run check     # typecheck + tests + docs/readme/examples staleness gates (the single gate)
```

`npm run provision:duckhts` installs the DuckHTS community extension for local `duckhts.read_bcf` tests.
`npm run provision:ducknng-owned` verifies the owned retry/auth/upload/TLS/RPC surface. Runtime Pi APIs are peer
dependencies supplied by Pi itself.

## Status & contributing

Pre-1.0 (`0.1.0`). The public API may still move. Issues and PRs welcome. Keep new
capability manifest/SQL-first: a new question should become data, a resolver adapter,
or SQL, not bespoke core code.

## License

[GPL-2.0-or-later](LICENSE) © sounkou-bioinfo
